---
title: "NGS Coverage Report"
author: "Carmelina Charalambous"
date: "11 March 2020"
output: html_document
editor options: 
 chunk_output_type: console
---



# Summary
This script creates an NGS coverage report to highlight genes with sub-optimal coverage (less than 30x coverage). This script can be be applied to any gene panel by replacing the file with your desired output file at line 60.


## Start
```{r start_section, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

#Time
Sys.time()

#Clean-up

rm(list=ls())
graphics.off()

#Set desired options
options(stringsAsFactors = F,
        warnPartialMatchArgs = T,
         warnPartialMatchAttr = T,
         warnPartialMatchDollar = T)


#Load necessery libraries
library(readr)
library(dplyr)
library(tidyr)
library(readxl)
library(xml2)
library(stringr)
library(openxlsx)
library(kableExtra)


```

```{r setwd, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

### Set working directory

setwd("/Volumes/SECURE/NGS_coverage")

```


### Load data and aggregate percentage coverage 30x by gene

```{r load_data, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

Raw_coverage_output <- read_excel("/Volumes/SECURE/NGS_coverage/coverage_30.xlsx")

Raw_coverage_output <- separate(Raw_coverage_output, "GeneSymbol;Accession", "GeneSymbol", sep = ";", remove = FALSE, convert = FALSE)


```

### Check if there is more than 1 gene accession number per gene
```{r accession_numbers, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

genes_inique <- unique(Raw_coverage_output$GeneSymbol)

accessions <- unique(Raw_coverage_output$`GeneSymbol;Accession`) 

# Each  gene was mapped to a unique accession number

kable(Raw_coverage_output[1:5,], format = "html", align = "c") %>%
  kable_styling(bootstrap_options = "condensed", full_width = T, position = "float_left", font_size = 10) %>%
  column_spec(1, bold = T, color = "#142851")

```


### Genes that have failed coverage at 30x are tabulated below
```{r less_than_100, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

# Calculate average percentage coverage per gene and also taking in consideration exon length (uniformity of coverage distribution decreases for longer exons)

aggregated_per_gene <- Raw_coverage_output %>%
  mutate(exon_length = abs(EndPosition - StartPosition)) %>% 
  group_by(GeneSymbol) %>%
  summarise(avg_percentage_30=weighted.mean(percentage30,exon_length)) %>% 
  filter(avg_percentage_30<100) %>% 
  select(GeneSymbol,avg_percentage_30)

kable(aggregated_per_gene, format = "html", align = "c") %>%
  kable_styling(bootstrap_options = "condensed", full_width = T, position = "float_left", font_size = 12) %>%
  column_spec(1, bold = T, color = "#142851")

```


### Clean_up
```{r final_section, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

ls()
Sys.time()

```




